---
title: "p8105_hw3_rt2640"
output: github_document
---

```{r loading libraries}
library(tidyverse)
library(p8105.datasets)
data("instacart")
data("brfss_smart2010")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

### Problem 1

The instacart dataset contains 1348617 observations and 15 variables. The dataframe provides a description of the orders on instacart including the items ordered, a description of their aisles and respective departments, along with a description of the user, the time of the order, days since the last order, and items reordered.

#### Problem 1 Part 1

```{r}
instacart_df = instacart %>%
  select(order_id:add_to_cart_order,aisle_id:department) %>% 
  group_by(aisle_id , aisle) %>%
  summarise(total = n()) %>%
  arrange(-total)
```
There are 134 aisles in total and most items are ordered from the fresh vegetables aisle 83 (n= 150609) followed by the fresh fruits aisle 24 (n =150473).

#### Problem 1 Part 2

```{r}
dept = instacart_df %>%
  filter(total > 10000) %>%
  arrange(-total)

dept %>%
  ggplot(aes(x = total, y = aisle)) +
  geom_point()

```

#### Problem 1 Part 3

```{r}
baking = instacart %>%
  group_by(aisle_id, aisle,product_name) %>%
  filter(aisle == "baking ingredients") %>%
  summarise(total = n()) %>%
  arrange(-total) %>%
  head(3)

dog_food = instacart %>%
  group_by(aisle_id, aisle,product_name) %>%
  filter(aisle == "dog food care") %>%
  summarise(total = n()) %>%
  arrange(-total) %>%
  head(3)

pack_veg_fruits = instacart %>%
  group_by(aisle_id, aisle,product_name) %>%
  filter(aisle == "packaged vegetables fruits") %>%
  summarise(total = n()) %>%
  arrange(-total) %>%
  head(3)

c1_table = bind_rows(baking, dog_food, pack_veg_fruits) %>%
  janitor::clean_names() %>%
  arrange(-total) %>% 
  as_tibble() %>% 
  knitr::kable()
c1_table

```

#### Problem 1 Part 4

```{r}
pla_table = instacart %>%
  group_by(product_name , order_dow) %>%
  filter(product_name == "Pink Lady Apples" | 
           product_name == "Coffee Ice Cream") %>%
  summarize(
    mean_hod = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hod
  ) %>%
  rename(
    Monday = "0",
    Tuesday = "1",
    Wednesday = "2",
    Thursday = "3",
    Friday = "4",
    Saturday = "5",
    Sunday = "6",
    Product = "product_name") %>%
  knitr::kable(digit = 2)
pla_table
```

### Problem 2

#### Problem 2 Part 1

```{r}
brfss_tidy = brfss_smart2010 %>%
  janitor::clean_names() %>%
  rename(state = locationabbr, location = locationdesc) %>%
  filter(topic == "Overall Health") %>%
  mutate(
    response = factor(response, levels = c("Poor","Fair","Good","Very good",
                                           "Excellent" )))
```

#### Problem 2 Part 2a

```{r}

brfss_2002 = brfss_tidy %>%
  filter(year == "2002") %>%
  group_by(state) %>%
  summarise(observed = n()/5) %>%
  filter(observed >= 7)
  

brfss_2010 = brfss_tidy %>%
  filter(year == "2010") %>%
  group_by(state) %>%
  summarise(observed = n()/5) %>%
  filter(observed >= 7)  
```

CT, FL, MA, NC, NJ, and PA are the states with 7 or more observation sites in 2002.
CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA are the states with 7 or more observation sites in 2010.

#### Problem 2 Part 2b

```{r}

excel_df = brfss_tidy %>%
  filter(response == "Excellent") %>%
  group_by(year,state) %>%
  summarise(
    avg_dv = mean(data_value, na.rm = TRUE)
  )

excel_df %>%
  ggplot(aes(x = year, y = avg_dv, color = state)) +
  geom_line(aes(group = state), alpha = .5) +
  labs(title = "Average values over time per State") +
  theme(legend.position = "none")
```

#### Problem 2 Part 2c

```{r}

nys_df = brfss_tidy %>%
  filter(state == "NY", year == "2006" | year == "2010") %>%
  group_by(response, year)

nys_df %>%
  ggplot(aes(x = response , y = data_value)) +
  geom_col() +
  facet_grid(~year) +
  labs(
    x = "Response",
    y = "Number of Responses",
    title = "Survey data in NYS State")
```
  
### Problem 3

#### Problem 3 Part 1 

```{r loading the dataset}
accel = read.csv(file = "./accel_data.csv")

accel_df = accel %>%
  janitor::clean_names() %>%
  mutate(
    weekday_vs_weekend = case_when(
      day == "Monday" ~ "Weekday",
      day == "Tuesday" ~ "Weekday",
      day == "Wednesday" ~ "Weekday",
      day == "Thursday" ~ "Weekday",
      day == "Friday" ~ "Weekday",
      day == "Saturday" ~ "Weekend",
      day == "Sunday" ~ "Weekend")) %>%
  rename(day_of_the_week = "day") %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_num",
    values_to = "activity_count")
```

The dataset contains 6 variables and 50400 observations. The variables included are the week, day of the week, whether it is a weekday or a weekend and activity counts per minute of everyday.


#### Problem 3 Part 2

```{r}
activity_day = accel_df %>%
  separate(activity_num, into = c("activity", "minute"), sep = "_") %>%
  group_by(day_id,day_of_the_week,week, weekday_vs_weekend) %>%
  summarize(
    total_activity = sum(activity_count)) %>%
  knitr::kable(digits = 4)

activity_day
```

The subject under observation overall has higher activity during the weekdays compared to the weekends on average across 5 weeks. He also has significantly lower total activity on Saturday of Week 4 and Week 5.

#### Problem 3 Part 3

```{r}

by_day = accel_df %>%
  separate(activity_num, into = c("activity", "minute"), sep = "_") %>%
  mutate(minute = as.numeric(minute)) %>%
  group_by(day_id,day_of_the_week,week, weekday_vs_weekend)

by_day %>%
  ggplot(aes(x = minute, y = activity_count , color = day_of_the_week)) +
  geom_point() +
  geom_line(alpha = .3) +
  theme(legend.position = "bottom") +
  viridis::scale_fill_viridis(discrete = TRUE)
```

The subject under observation has an overall lower activity on tuesdays across all 5 weeks of observation. The average activity between 0-400 minutes (night time) is lower across all days. The activity trend on Fridays is on average higher during the evening hours.

  


  

           
  
  


  

  



