---
title: "p8105_hw3_rt2640"
output: github_document
---

```{r loading libraries}
library(tidyverse)
library(p8105.datasets)
data("instacart")
data("brfss_smart2010")

```

### Problem 1

The instacart dataset contains 1348617 observations and 15 variables. The dataframe provides a description of the orders on instacart including the items ordered, a description of their aisles and respective departments, along with a description of the user, the time of the order, days since the last order, and items reordered.

#### Problem 1 Part 1

```{r}
instacart_df = instacart %>%
  select(order_id:add_to_cart_order,aisle_id:department) %>% 
  group_by(aisle_id , aisle) %>%
  summarise(total = n()) %>%
  arrange(-total)
```
There are 134 aisles in total and most items are ordered from the fresh vegetables aisle 83 (n= 150609) followed by the fresh fruits aisle 24 (n =150473).

#### Problem 1 Part 2

```{r}
dept = instacart_df %>%
  filter(total > 10000) %>%
  arrange(-total)

dept %>%
  ggplot(aes(x = total, y = aisle)) +
  geom_point()

```

#### Problem 1 Part 3

```{r}
baking = instacart %>%
  group_by(aisle_id, aisle,product_name) %>%
  filter(aisle == "baking ingredients") %>%
  summarise(total = n()) %>%
  arrange(-total) %>%
  head(3)

dog_food = instacart %>%
  group_by(aisle_id, aisle,product_name) %>%
  filter(aisle == "dog food care") %>%
  summarise(total = n()) %>%
  arrange(-total) %>%
  head(3)

pack_veg_fruits = instacart %>%
  group_by(aisle_id, aisle,product_name) %>%
  filter(aisle == "packaged vegetables fruits") %>%
  summarise(total = n()) %>%
  arrange(-total) %>%
  head(3)

c1_table = bind_rows(baking, dog_food, pack_veg_fruits) %>%
  janitor::clean_names() %>%
  arrange(-total) %>% 
  as_tibble() %>% 
  knitr::kable()
c1_table

```

#### Problem 1 Part 4

```{r}
pla_table = instacart %>%
  group_by(product_name , order_dow) %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  summarize(
    mean_hod = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hod
  ) %>%
  rename(
    Monday = "0",
    Tuesday = "1",
    Wednesday = "2",
    Thursday = "3",
    Friday = "4",
    Saturday = "5",
    Sunday = "6",
    Product = "product_name") %>%
  knitr::kable(digit = 2)
pla_table
```

### Problem 2

#### Problem 2 Part 1

```{r}
brfss_tidy = brfss_smart2010 %>%
  janitor::clean_names() %>%
  rename(state = locationabbr, location = locationdesc) %>%
  filter(topic == "Overall Health") %>%
  mutate(
    response = factor(response, levels = c("Poor","Fair","Good","Very good","Excellent" )))
```
  






